### 2023-04-17 output/presentations/cep_internal_230419/presentation.qmd
Kept working on all data requests. Started writing presentation for internal CEP meeting.

---
title: "CEP Group Meeting"
date: 2023-04-19
author: Lorenzo Fabbri
format:
  revealjs:
    footer: "Lorenzo Fabbri"
    toc: true
    toc_float: true
    toc-depth: 2
    toc-location: body
link-citations: true
linkcolor: blue
execute:
  echo: false
  warning: false
  error: false
  message: false
editor_options: 
  chunk_output_type: console
---

# Overview

## The Research Questions

# Triangulation

# Reliability and Replicability