### 2023-05-22 docs/diary.qmd
Updated diary. Error when rendering causal roadmap.

---
title: "Diary"
author: "Lorenzo Fabbri"
date-format: "DD/MM/YYYY"
format:
  pdf
editor_options: 
  chunk_output_type: console
---

# 28/05/2023

Remaining doubts:

* Should I include the covariates when imputing the exposures?
* Should I impute the covariates by cohort?

## Missing values by variables

* There is a clear pattern for some covariates and some cohorts.
* Possible solutions:
  * Drop the variables with a percentage of missing values greater than a pre-specified threshold, by cohort. Perform analyses by cohort, adjusting for different covariates, and pool the results.
  * Drop the variables with a percentage of missing values greater than a pre-specified threshold, by cohort and overall. Perform a single analysis. Then, perform sensitivity analyses by keeping those variables and dropping the cohorts with a *high* fraction of missing values.

## Missing values by cases

* `KAN` participants seem to be the ones with the largest fraction of missing values.
* Possible solutions:
  * Drop these cases and perform the analyses. Then perform sensitivity analysis by including these cases and imputing the missing values.