### 2023-07-13 output/presentations/cep_internal_230726/presentation.qmd
Started presentation internal meeting. Reformatted parameters for analyses.

---
title: "CEP Group Meeting"
date: 2023-07-26
author: Lorenzo Fabbri
institute: ISGlobal

format:
  beamer:
    pdf-engine: xelatex
    fontsize: "9pt"
    toc: false
    toc-depth: 2

slide-number: true

bibliography: ../../../../../PhD.bib
link-citations: true
linkcolor: blue

execute:
  echo: false
  warning: false
  error: false
  message: false

editor_options: 
  chunk_output_type: console
---

```{r}
#| include: false
here::i_am("presentation.qmd")
```

## Overview

- **Aim**: to research the *short-term* effects of postnatal exposures to non-persistent EDCs on neurodevelopment and neurobehavior in childhood, and how the metabolome and the proteome might *mediate* these effects.
- **How**: by making use of the principles and criteria of triangulation. We will primarily rely on the use of MTPs in combination with TMLE.
- **Population**: the HELIX sub-cohort, consisting of $\text{N}=1200$ mother-child pairs.
- **Exposures**: non-persistent EDCs (phenols, phthalates, and organophosphate compounds), measured in childhood in a pool of two urine samples.
- **Outcomes**
  - Raven’s Coloured Progressive Matrices, for assessing non-verbal intelligence.
  - Computerised n-back test, for assessing working memory.
  - Attention Network Test, to provide a measure of the efficiency of three different functions of attention.
  - Child Behavior Checklist, for assessing behavioural and emotional problems.
- ***Mediators***: serum and urine metabolome, plasma proteome.

## The Research Questions

```{mermaid}
%%| fig-cap: Simplified DAG.
%%| label: fig-diag-dag
%%| fig-align: center
flowchart LR
  exposure(chemical) --> mediator(metabolite)
  mediator --> outcome(neurodevelopment)
  exposure --> outcome
```

## Triangulation

```{mermaid}
%%| fig-cap: Diagram summarizing research questions and methods.
%%| label: fig-diagram
%%| fig-align: center
%%| fig-height: 3
flowchart LR
  paper[Paper 3] --> rq1("RQ1
                          neuro~chemical")
  paper          --> rq2("RQ2
                          met~chemical")
  paper          --> rq3("RQ3
                          neuro~met")
  rq1            --> rqReg("Observational study.
                            Method: Regression with MTPs.
                            Causal estimand: phi.
                            Uncertainty quantification: bootstrapping.")
  rq1            --> rq1outNeg("Outcome negative control.
                                Method: Regression with GLMs/GAMs.
                                Uncertainty quantification: bootstrapping.")
  rq1            --> rqLit("Literature search")

  rq2            --> rqReg
  rq2            --> rqCrossCohort("Cross-cohort comparison.
                                    Method: Regression with GLMs/GAMs.
                                    Uncertainty quantification: bootstrapping.")
  rq2            --> rqLit

  rq3            --> rqReg
  rq3            --> rqCrossCohort
  rq3            --> rq3mr("Mendelian Randomization.
                            Causal estimand: .")
```