### 2023-07-21 output/presentations/cep_internal_230726/presentation.qmd
Added DAG for RQ2, and corresponding file. Worked more on presentation for group meeting.

---
title: "CEP Group Meeting"
subtitle: "Update on *paper 3* of my PhD"
date: 2023-07-26
author: Lorenzo Fabbri
institute: ISGlobal

format:
  beamer:
    pdf-engine: xelatex
    fontsize: "9pt"
    toc: false
    toc-depth: 2
geometry:
  - left=10mm
  - right=10mm

slide-number: true

bibliography: ../../../../../PhD.bib
link-citations: true
linkcolor: blue

execute:
  echo: false
  warning: false
  error: false
  message: false

editor_options: 
  chunk_output_type: console
---

```{r}
#| include: false
here::i_am("presentation.qmd")
```

## Overview

- **Aim**: to research the *short-term* effects of childhood exposures to non-persistent EDCs on neurodevelopment in childhood, and how the metabolome **might** *mediate* these effects.
- **How**: by making use of the principles and criteria of triangulation.
- **Population**: the HELIX sub-cohort, consisting of $\text{N}=1200$ mother-child pairs.
- **Exposures**: non-persistent EDCs (phenols, phthalates, and organophosphate compounds), measured in childhood in a pool of two urine samples.
- **Outcome**: Attention Network Test, to provide a measure of the efficiency of three different functions of attention.
- ***Mediators***: urine metabolites (focus on corticosteroids).

## The Research Questions

```{mermaid}
%%| fig-cap: Simplified *DAG*.
%%| label: fig-diag-dag
%%| fig-align: center
flowchart LR
  exposure(chemical) --> mediator(metabolite)
  mediator --> outcome(neurodevelopment)
  exposure --> outcome
```

Issues and limitations:

- The effect of exposure to these chemicals on neurodevelopment has been studied already, although there are clear issues with that study (e.g., wrong model, DAG **not** tailored to the exposures of interest).
- The effect of exposure to these chemicals on these **newly** measured metabolites could be identifiable, although they were measured in the same samples. This question poses some challenges since they were all measured in urine (inter-individual variability in urine dilution).
- The effect of these **newly** measured metabolites on neurodevelopment could be identifiable, since they were measure in the urine samples (\textcolor{red}{which aliquot?}) collected before the visit.

## Triangulation

```{mermaid}
%%| fig-cap: Diagram summarizing research questions and methods.
%%| label: fig-diagram
%%| fig-align: center
%%| fig-height: 3
flowchart LR
  paper[Paper 3] --> rq1("RQ1
                          neuro~chemical")
  paper          --> rq2("RQ2
                          met~chemical")
  paper          --> rq3("RQ3
                          neuro~met")
  rq1            --> rqReg("Observational study.
                            Method: Regression with MTPs.
                            Causal estimand: phi.
                            Uncertainty quantification: bootstrapping.")
  rq1            --> rq1outNeg("Outcome negative control.
                                Method: Regression with GLMs/GAMs.
                                Uncertainty quantification: bootstrapping.")
  rq1            --> rqLit("Literature search")

  rq2            --> rqReg
  rq2            --> rqCrossCohort("Cross-cohort comparison.
                                    Method: Regression with GLMs/GAMs.
                                    Uncertainty quantification: bootstrapping.")
  rq2            --> rqLit

  rq3            --> rqReg
  rq3            --> rqCrossCohort
  rq3            --> rq3mr("Mendelian Randomization.
                            Causal estimand: .")
  style rq2 fill:#7faac2
  style rq3 fill:#7faac2
```

## `myphd` R package

![`myphd`](artwork/myphd.png){width=300}

- URL manual: https://isglobal-cep.github.io/myphd/
- To install it: `remotes::install_github("isglobal-cep/myphd")`

:::: {.columns align=center}

::: {.column width="33%" align=top}
```{r, eval=FALSE, echo=TRUE}
myphd::create_formula(
  dat,
  outcome,
  exposure,
  covariates,
  method,
  add_inter_exposure,
  add_splines_exposure,
  df_splines,
  threshold_smooth,
  threshold_k
)
```
:::

::: {.column width="33%" align=top}
```{r, eval=FALSE, echo=TRUE}
myphd::describe_data(
  dat,
  id_var,
  grouping_var
)
```
:::

::: {.column width="33%" align=top}
```{r, eval=FALSE, echo=TRUE}
myphd::preproc_data(
  dat,
  outcome = NULL,
  dic_steps,
  id_var,
  by_var
)
```
:::

::::

## Current setup

Variables:

- Clinical outcome: Attention Network Test (`hs_hitrtse`, Hit Reaction Time Standard Error).

Methods:

- Weights estimation (`chemical ~ covariates`): energy balancing.
- Effect estimation (`outcome ~ chemical + covariates`): `glm` with natural splines for exposure.

## Population description: correlations

![`Correlation numerical variables.`](artwork/corr.png)

## Preliminary results

## Next steps