### 2023-08-02 output/paper/sections/methods.qmd
Using targets to build paper. Checking that labels levels factors are displayed in summary tables.

## Study population {#sec-study-population}

## Exposure assessment {#sec-exp-assessment}

## Metabolomics {#sec-metabolomics}

## Neurodevelopment {#sec-neurodevelopment}

## Covariates {#sec-covars}

## Statistical analysis {#sec-stat-analysis}