### 2023-08-09 output/paper/sections/discussion.qmd
Added function to plot densities of variables by grouping variable and with CDF. Working on avg_comparisons with  term and with functionality to compute quantiles by grouping variables since for instance chemical levels vary a lot by cohort.

List of possible limitations:

* The percentiles of the exposures might differ considerably by cohort. When using `avg_comparisons`, we do not consider this. This might cause cohort-dependent positivity issues (?).