### 2023-09-08 output/paper/figures.qmd
Added details for paper if published in EHP. Still no line numbering in DOCX nor PDF.

---
title: "Figures"

format:
  pdf:
    pdf-engine: lualatex
    toc: false
    include-in-header:
      text: |
        \usepackage{rotating}
        \usepackage{float}
        \usepackage{typearea}

link-citations: true
linkcolor: blue

execute:
  echo: false
  warning: false
  error: false
  message: false
editor_options: 
  chunk_output_type: console
---

- Number figures according to the order in which they are first mentioned in the main text.
- Clearly label all axes, giving both the measure and the unit of measurement.
- Not larger than 7.5 inches in width.
- Save and submit each main text figure as an individual file (PDF with fonts embedded). Include the figure number in the file name of each figure.
- Provide main text figure captions on a new page of the main text after the tables.
- Include a title for the entire figure and descriptors for each panel.

# CTD Results
```{r}
#| label: fig-ctd-res
#| fig.cap: Barplot CTD.
#| fig.height: 10
#| fig.width: 10

path <- "results/files/CTD_chem_diseases_1680702221550_05042023.csv"
ctd <- myphd::load_ctd(path = path, 
                       filter_evidence = TRUE)
myphd::plot_ctd(ctd)
```