### 2023-09-08 output/paper/sections/discussion.qmd

## Key results {#sec-disc-res-key}

- Summarise key results with reference to study objectives.
- Provide a review of the relevant literature to put the study findings into context.
  - It should be complete and balanced, including inconsistent results.
  - It should include, for each source, sufficient details: study design, sample size, population, specific exposures and outcomes.

## Limitations {#sec-limitations}

- Discuss limitations of the study, taking into account sources of potential bias or imprecision.
- Discuss both direction and magnitude of any potential bias.

## Interpretation {#sec-interpretation}

- End with a summary of the key findings and their implications for the study hypotheses, future research, and policy.
- Give a cautious overall interpretation of results considering objectives, limitations, multiplicity of analyses, results from similar studies, and other relevant evidence.

## Generalisability {#sec-general}

- Discuss the generalisability (external validity) of the study results.