### 2023-09-08 output/paper/sections/introduction.qmd

## Background and rationale {#sec-background}

- Brief review of the literature to summarize current knowledge.
  - Acknowledge inconsistencies.
  - For each study, indicate whether it was observational or experimental, and note key characteristics of study populations or experimental models.
- Explain the scientific background and rationale for the investigation being reported.
  - Identify knowledge gaps addressed by the current study.
- Provide context for the study: include information on exposures and outcomes, and why they are relevant to environmental health.

## Objectives {#sec-objectives}

- Provide a clear description of the study hypotheses/aims/objectives, and eventually an overview of the approach used to address them.