### 2023-09-08 output/paper/sections/methods.qmd

## Study design {#sec-design}

- Present key elements of study design

## Setting {#sec-setting}

- Describe the setting, locations, and relevant dates, including periods of recruitment, exposure, follow-up, and data collection.

## Participants {#sec-participants}

- Cohort study: eligibility criteria, and the sources and methods of selection of participants. Describe methods of follow-up.
- Cross-sectional study: give eligibility criteria, and the sources and methods of selection of participants.
- Describe informed consent protocols.
- Report how and by whom *race* or *ethnicity* was defined, and why this information was included in the study design. Disaggregate race and ethnicity data to the fullest extent possible.

## Variables {#sec-vars}

- Clearly define all outcomes, exposures, predictors, potential confounders, and effect modifiers.
- Explain the rationale for treating race as an exposure, confounder, effect modifier, or other type of variable in analyses.

### Confounders {#sec-confounders}

### Endocrine disrupting chemicals {#sec-edcs}

### Corticosteroids {#sec-steroids}

### Neurodevelopment {#sec-neurodevelopment}

## Data sources and measurement {#sec-dat-sources}

- For each variable of interest, give sources of data and details of methods of assessment (measurement).

## Bias {#sec-bias}

- Describe any efforts to address potential sources of bias.

## Study size {#sec-size}

- Explain how the study size was arrived at.

## Quantitative variables {#sec-quant-vars}

- Explain how quantitative variables were handled in the analyses. If applicable, describe which groupings were chosen and why.

## Statistical methods {#sec-stat-methods}

- Methods for selecting potential confounders (provide DAGs).
- Describe all statistical methods with assumptions, including those used to control for confounding.
- Describe any methods used to examine subgroups and interactions.
- Explain how missing data were addressed.
- Cohort study: explain how loss to follow-up was addressed.
- Cross-sectional study: describe analytical methods taking account of sampling strategy.
- Describe any sensitivity analyses.
- When referring to previous publications for methods' details, include a brief description of the approach, key assumptions and limitations, and any deviation.
- Names and version numbers for the used software packages, including non-data arguments if deviating from the default ones.