### 2023-09-08 output/paper/si.qmd

---
title: "Supplementary Material"

format:
  docx:
    reference-doc: reference_document.docx

bibliography: bibliography.bib
csl: ama11.csl

link-citations: true
linkcolor: blue

execute:
  echo: false
  warning: false
  error: false
  message: false
editor_options: 
  chunk_output_type: console
---

- Should be limited to results of secondary analyses and background details needed to ensure transparency:
  - Questionnaires.
  - Tables and figures with results of sensitivity analyses.
  - Tables with numeric data corresponding to results shown in the main text figures.
  - DAGs used to select model covariates.
- Text description of study methods should NOT go in the SM.
- Order the content: text, supplemental tables, and supplemental figures.
- Provide the title for each figure below the figure on the same manuscript page as the figure itself.
- Provide the title for each table above the table and the notes below the table.
- Number supplemental tables and figures separately according to the order in which they are first mentioned in the main text, using an "S" prefix.
- Conclude the SM with a list of references, even if they are also cited in the main paper.