### 2023-09-08 output/paper/tables.qmd

---
title: "Tables"

format:
  docx:
    reference-doc: reference_document.docx

link-citations: true
linkcolor: blue

execute:
  echo: false
  warning: false
  error: false
  message: false
editor_options: 
  chunk_output_type: console
---

- Begin each table on a new page after the list of references.
- Tables may not contain parts.
- List footnotes after the general one, one per line and indicated with a lowercase italicized letter.

# Study populations
```{r}

```

# Exposures
```{r}
#| label: tbl-chem-info
#| tbl-cap: Information about the non-persistent EDCs.

edcs_info <- myphd::edcs_information() |>
  tidylog::select(-c(chem_id, smiles))
gt::gt(edcs_info, groupname_col = "class") |>
  gt::cols_label(
    full_name = "Compound", 
    short_name = "Symbol", 
    pubchem_cid = "PubChem CID", 
    ctd_id = "CTD ID", 
    exposome_explorer_id = "Exposome Explore ID", 
    parental_compound = "Parental compound"
  ) |>
  gt::tab_style(
    locations = gt::cells_column_labels(columns = everything()), 
    style = list(
      gt::cell_borders(sides = "bottom", weight = gt::px(3)), 
      gt::cell_text(weight = "bold")
    )
  ) |>
  gt::tab_style(
    locations = gt::cells_row_groups(group = dplyr::everything()), 
    style = list(gt::cell_text(weight = "bold"))
  )
```