### 2023-09-13 output/paper/figures.qmd
Working on structure of paper. Issues with cross-referencing tables in Quarto.

---
title: "Figures"

format:
  pdf:
    pdf-engine: lualatex
    number-sections: true
    toc: false
    include-in-header:
      text: |
        \usepackage{lineno}
        \usepackage{rotating}
        \usepackage{float}
        \usepackage{typearea}
    include-before-body:
      text: |
        \linenumbers

link-citations: true
linkcolor: blue

editor_options: 
  chunk_output_type: console

execute:
  echo: false
  warning: false
  error: false
  message: false
---

- Number figures according to the order in which they are first mentioned in the main text.
- Clearly label all axes, giving both the measure and the unit of measurement.
- Not larger than 7.5 inches in width.
- Save and submit each main text figure as an individual file (PDF with fonts embedded). Include the figure number in the file name of each figure.
- Provide main text figure captions on a new page of the main text after the tables.
- Provide the title for each figure below the figure on the same manuscript page as the figure itself.
- Include a title for the entire figure and descriptors for each panel.

# Participants

- Flow diagram: number of individuals at each stage of study. Report numbers potentially eligible, examined for eligibility, confirmed eligible, included in the study, completing follow-up, and analysed.

# Main results

- Forest plot for RQ2 (group by chemical class, and for each chemical group bars for metabolites) and RQ3.