### 2023-09-13 output/paper/tables.qmd

---
title: "Tables"

format:
  docx:
    reference-doc: reference_document.docx

tbl-cap-location: top
link-citations: true
linkcolor: blue

editor_options: 
  chunk_output_type: console

execute:
  echo: false
  warning: false
  error: false
  message: false
---

- Begin each table on a new page after the list of references.
- Tables may not contain parts.
- Provide the title for each table above the table and the notes below the table.
- List footnotes after the general one, one per line and indicated with a lowercase italicized letter.

# Descriptive data

## Study populations

- Basic characteristics of study participants (demographic, clinical, social), by cohort and overall.
  - Include all covariates included in primary or secondary analyses.
  - Indicate number of participants with missing data for each variable of interest.
  - Report numbers of outcome events or summary measures.

## Endocrine disruptors

- Levels of unprocessed biomarkers (chemicals), with IDs.
- Report: minimum and maximum, percentiles, and number of observations above/below the assay's LOD and LOQ.

```{r}
#| label: tbl-chem-info
#| tbl-cap: "Information about the non-persistent EDCs."

edcs_info <- myphd::edcs_information() |>
  tidylog::select(-c(chem_id, smiles))
gt::gt(edcs_info,
       groupname_col = "class") |>
  gt::tab_header(
    title = "Information about the non-persistent EDCs."
  ) |>
  gt::cols_label(
    full_name = "Compound",
    short_name = "Symbol",
    pubchem_cid = "PubChem CID",
    ctd_id = "CTD ID",
    exposome_explorer_id = "Exposome Explore ID",
    parental_compound = "Parental compound"
  ) |>
  gt::tab_style(
    locations = gt::cells_column_labels(columns = everything()),
    style = list(
      gt::cell_borders(sides = "bottom", weight = gt::px(3)),
      gt::cell_text(weight = "bold")
    )
  ) |>
  gt::tab_style(
    locations = gt::cells_row_groups(group = dplyr::everything()),
    style = list(gt::cell_text(weight = "bold"))
  )
```

## Corticosteroids

- Levels of unprocessed biomarkers (metabolites), with IDs.

# Main results

- RQ2: group chemicals by class, and for each chemical report results with metabolites (`avg_comparisons` with estimate, SE, s-value, CI).
- RQ3: report results with outcome (`avg_comparisons` with estimate, SE, s-value, CI).

# Other analyses