### 2023-11-07 output/paper/figures.qmd
Modified pipeline so that targets for processing results do not actually produce eg plots but only ready to use tibbles for plots and tables. Tested pipeline to create paper but I think I still have to manually reference tables and figures in manuscript.

---
title: "Figures"
format:
  pdf:
    pdf-engine: lualatex
    number-sections: true
    toc: false
    include-in-header:
      text: |
        \usepackage{lineno}
        \usepackage{rotating}
        \usepackage{float}
        \usepackage{typearea}
    include-before-body:
      text: |
        \linenumbers
link-citations: true
linkcolor: blue
editor_options: 
  chunk_output_type: console
execute:
  echo: false
  warning: false
  error: false
  message: false
---

```{r}
source("code/proc_res.R")
```

```{r vars}
store <- ifelse(
  Sys.getenv("is_hpc"), 
  paste0(
    "/PROJECTES/HELIX_OMICS/DATA_PREVIOUS_MIGRATION/",
    "lorenzoF/data/data_paper3/_targets/_targets_res"
  ), 
  "~/mounts/rstudioserver/data/data_paper3/_targets/_targets_res"
)

num_digits_est = 3
num_digits_sig = 2
```

- Number figures according to the order in which they are first mentioned in the main text.
- Clearly label all axes, giving both the measure and the unit of measurement.
- Smallest type size: 8pt.
- Maximum width: 7 inches in portrait, 9 inches in landscape mode.
- Multi-panel figures: capital letters followed by closing parenthesis, completely above and to the left of the panel.
- Use log-scale when plotting relative measures of effects.
- Figure titles and legends should be submitted as single paragraph.
- The text of figure legends typed after the reference list.
- Save and submit each main text figure as an individual file (PDF with fonts embedded).

# Participants

- Flow diagram: number of individuals at each stage of study. Report numbers potentially eligible, examined for eligibility, confirmed eligible, included in the study, completing follow-up, and analysed.

# Main results

- Forest plot for RQ2 (group by chemical class, and for each chemical group bars for metabolites) and RQ3.

## Marginal comparisons

```{r}
targets::tar_load(
  dplyr::matches("marginal_comparisons_(\\d)$"),
  store = store
)
```

```{r marginal-1}
#| label: fig-marginal-1
#| fig-cap: Average comparisons
#| fig-width: 6
#| fig-height: 7

tidy_res_meffects(
  df = marginal_comparisons_1,
  sa_var = NULL,
  which_res = "comparisons",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$plot
```

```{r marginal-2}
#| label: fig-marginal-2
#| fig-cap: Average comparisons
#| fig-width: 10
#| fig-height: 10

tidy_res_meffects(
  df = marginal_comparisons_2,
  sa_var = NULL,
  which_res = "comparisons",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$plot
```

```{r marginal-3}
#| label: fig-marginal-3
#| fig-cap: Average comparisons
#| fig-width: 5
#| fig-height: 5

tidy_res_meffects(
  df = marginal_comparisons_3,
  sa_var = NULL,
  which_res = "comparisons",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$plot
```