### 2023-11-07 output/paper/si.qmd

---
title: "Supplementary Material"

format:
  docx:
    reference-doc: reference_document.docx

crossref:
  fig-title: Supplementary Figure
  tbl-title: Supplementary Table

bibliography: bibliography.bib
csl: ama11.csl

link-citations: true
linkcolor: blue

editor_options: 
  chunk_output_type: console

execute:
  echo: false
  warning: false
  error: false
  message: false
---

```{r}
source("code/proc_res.R")
```

```{r vars}
store <- ifelse(
  Sys.getenv("is_hpc"), 
  paste0(
    "/PROJECTES/HELIX_OMICS/DATA_PREVIOUS_MIGRATION/",
    "lorenzoF/data/data_paper3/_targets/_targets_res"
  ), 
  "~/mounts/rstudioserver/data/data_paper3/_targets/_targets_res"
)

num_digits_est = 3
num_digits_sig = 2
```

- Should be limited to results of secondary analyses and background details needed to ensure transparency:
  - Questionnaires.
  - Tables and figures with results of sensitivity analyses.
  - Tables with numeric data corresponding to results shown in the main text figures.
  - DAGs used to select model covariates.
- Text description of study methods should NOT go in the SM.
- Number supplemental tables and figures separately according to the order in which they are first mentioned in the main text, using an "S" prefix.

# Supplementary information

## Directed Acyclic Graphs

```{r dag-1}
targets::tar_load("all_dags", store = store)
cat(all_dags$chem_to_out)
```

```{r dag-2}
cat(all_dags$chem_to_marker)
```

```{r dag-3}
cat(all_dags$marker_to_out)
```

# Supplementary tables

- Provide the title for each table above the table and the notes below the table.

## Descriptive data

### Study populations

- Detailed characteristics of study participants (demographic, clinical, social), by cohort and overall.

```{r codebook-1}
#| label: tbl-codebook-1
#| tbl-cap: Codebook for covariates used in RQ1

targets::tar_load(
  dplyr::contains("codebook_"),
  store = store
)
codebook_1
```

```{r codebook-2}
#| label: tbl-codebook-2
#| tbl-cap: Codebook for covariates used in RQ2

codebook_2
```

```{r codebook-3}
#| label: tbl-codebook-3
#| tbl-cap: Codebook for covariates used in RQ3

codebook_3
```

## Other results

### Balancing weights for effect modification

```{r}
targets::tar_load(
  dplyr::matches("weights_(\\d)SA$"),
  store = store
)
```

```{r weights-1sa}
#| label: tbl-weights-1sa
#| tbl-cap: Weights

tidy_res_weighted_fits(
  dat_tbl = weights_1SA$dat_tbl,
  dat_plt = weights_1SA$dat_plt,
  sa_var = "e3_sex"
)$table
```

```{r weights-2sa}
#| label: tbl-weights-2sa
#| tbl-cap: Weights

tidy_res_weighted_fits(
  dat_tbl = weights_2SA$dat_tbl,
  dat_plt = weights_2SA$dat_plt,
  sa_var = "e3_sex"
)$table
```

```{r weights-3sa}
#| label: tbl-weights-3sa
#| tbl-cap: Weights

tidy_res_weighted_fits(
  dat_tbl = weights_3SA$dat_tbl,
  dat_plt = weights_3SA$dat_plt,
  sa_var = "e3_sex"
)$table
```

# Supplementary figures

- Provide the title for each figure below the figure on the same manuscript page as the figure itself.

## Main results

- ADRF and AMEF for RQ2 and RQ3.
- Whole distribution (i.e., density) of e.g., risk differences (instead of marginalized estimates), with lines for mean and median.

## Other results

### Marginal comparisons for effect modification

```{r}
targets::tar_load(
  dplyr::matches("marginal_comparisons_(\\d)SA$"),
  store = store
)
```

```{r marginal-1sa}
#| label: fig-marginal-1sa
#| fig-cap: Average comparisons for effect modification
#| fig-width: 6
#| fig-height: 7

tidy_res_meffects(
  df = marginal_comparisons_1SA,
  sa_var = "e3_sex",
  which_res = "comparisons",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$plot
```

```{r marginal-2sa}
#| label: fig-marginal-2sa
#| fig-cap: Average comparisons for effect modification
#| fig-width: 10
#| fig-height: 10

tidy_res_meffects(
  df = marginal_comparisons_2SA,
  sa_var = "e3_sex",
  which_res = "comparisons",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$plot
```

```{r marginal-3sa}
#| label: fig-marginal-3sa
#| fig-cap: Average comparisons for effect modification
#| fig-width: 5
#| fig-height: 5

tidy_res_meffects(
  df = marginal_comparisons_3SA,
  sa_var = "e3_sex",
  which_res = "comparisons",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$plot
```

# References {.unnumbered}

::: {#refs}
:::