### 2023-11-09 output/paper/sections/discussion.qmd
Kept working on processing results for paper.

## Key results {#sec-disc-res-key}

- Summarise key results with reference to study objectives.
- Provide a review of the relevant literature to put the study findings into context.
  - It should be complete and balanced, including inconsistent results.
  - It should include, for each source, sufficient details: study design, sample size, population, specific exposures and outcomes.

### What does the literature say?

- EDCs and neurodevelopment (ANT).
- EDCs and corticosteroids.
- Corticosteroids and neurodevelopment (ANT).

## Limitations {#sec-limitations}

- Discuss limitations of the study, taking into account sources of potential bias or imprecision.
- Discuss both direction and magnitude of any potential bias.

Some limitations:

- Cross-sectional study.
- Chemicals measured in night and morning samples, whereas metabolites (the outcome) were measured only in night samples.
- Cortisol measured at night, when should be lowest.
- Change of estimand when trimming weights.
- Model misspecification.
- Mixtures effect.
- Residual confounding.
- Some confounders were not used since large percentage of missing values.
- Multiple comparisons.

## Interpretation {#sec-interpretation}

- End with a summary of the key findings and their implications for the study hypotheses, future research, and policy.
- Give a cautious overall interpretation of results considering objectives, limitations, multiplicity of analyses, results from similar studies, and other relevant evidence.

## Generalisability {#sec-general}

- Discuss the generalisability (external validity) of the study results.