### 2023-11-09 output/paper/tables.qmd

---
title: "Tables"

format:
  docx:
    reference-doc: reference_document.docx

link-citations: true
linkcolor: blue

editor_options:
  chunk_output_type: console

execute:
  echo: false
  warning: false
  error: false
  message: false
---

```{r}
source("code/proc_res.R")
```

```{r vars}
store <- ifelse(
  Sys.getenv("is_hpc"), 
  paste0(
    "/PROJECTES/HELIX_OMICS/DATA_PREVIOUS_MIGRATION/",
    "lorenzoF/data/data_paper3/_targets/_targets_res"
  ), 
  "~/mounts/rstudioserver/data/data_paper3/_targets/_targets_res"
)

num_digits_est = 3
num_digits_sig = 2
```

# Descriptive data

## Study populations

```{r pop-desc}
#| label: tbl-pop-desc
#| tbl-cap: Population description

targets::tar_load("table1", store = store)
table1
```

## Endocrine disruptors

```{r edc-desc}
#| label: tbl-edc-desc
#| tbl-cap: Levels of unprocessed chemicals

targets::tar_load("desc_chems_mets", store = store)
desc_chems_mets$desc_chems
```

## Corticosteroids

```{r met-desc}
#| label: tbl-met-desc
#| tbl-cap: Levels of unprocessed corticosteroids

desc_chems_mets$desc_mets
```

# Main results

## Balancing weights

```{r}
targets::tar_load(
  dplyr::matches("balance_(\\d)$"),
  store = store
)
```

```{r balance-1}
balance_1 |>
  gt::gt() |>
  gt::fmt_number()
```

```{r}
targets::tar_load(
  dplyr::matches("weights_(\\d)$"),
  store = store
)
```

```{r weights-1}
#| label: tbl-weights-1
#| tbl-cap: Weights

tidy_res_weighted_fits(
  dat_tbl = weights_1$dat_tbl,
  dat_plt = weights_1$dat_plt,
  sa_var = NULL
)$table
```

```{r weights-2}
#| label: tbl-weights-2
#| tbl-cap: Weights

tidy_res_weighted_fits(
  dat_tbl = weights_2$dat_tbl,
  dat_plt = weights_2$dat_plt,
  sa_var = NULL
)$table
```

```{r weights-3}
#| label: tbl-weights-3
#| tbl-cap: Weights

tidy_res_weighted_fits(
  dat_tbl = weights_3$dat_tbl,
  dat_plt = weights_3$dat_plt,
  sa_var = NULL
)$table
```

## Marginal comparisons

```{r}
targets::tar_load(
  dplyr::matches("marginal_comparisons_(\\d)$"),
  store = store
)
```

```{r marginal-1}
#| label: tbl-marginal-1
#| tbl-cap: Average comparisons

tidy_res_meffects(
  df = marginal_comparisons_1,
  sa_var = NULL,
  which_res = "comparisons",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$table
```

```{r marginal-2}
#| label: tbl-marginal-2
#| tbl-cap: Average comparisons

tidy_res_meffects(
  df = marginal_comparisons_2,
  sa_var = NULL,
  which_res = "comparisons",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$table
```

```{r marginal-3}
#| label: tbl-marginal-3
#| tbl-cap: Average comparisons

tidy_res_meffects(
  df = marginal_comparisons_3,
  sa_var = NULL,
  which_res = "comparisons",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$table
```

# Other analyses

## Marginal hypothesis

```{r}
targets::tar_load(
  dplyr::matches("marginal_hypothesis_(\\d)SA$"),
  store = store
)
```

```{r hypothesis-1}
#| label: tbl-hypothesis-1
#| tbl-cap: Hypothesis testing for effect modification

tidy_res_meffects(
  df = marginal_hypothesis_1SA,
  sa_var = "e3sex",
  which_res = "hypothesis",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$table
```

```{r hypothesis-2}
#| label: tbl-hypothesis-2
#| tbl-cap: Hypothesis testing for effect modification

tidy_res_meffects(
  df = marginal_hypothesis_2SA,
  sa_var = "e3sex",
  which_res = "hypothesis",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$table
```

```{r hypothesis-3}
#| label: tbl-hypothesis-3
#| tbl-cap: Hypothesis testing for effect modification

tidy_res_meffects(
  df = marginal_hypothesis_3SA,
  sa_var = "e3sex",
  which_res = "hypothesis",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$table
```