### 2023-11-14 output/paper/sections/introduction.qmd
Added table LLOQs for steroids. Started kinda intro to paper.

In the present study, we estimated associations between 1) non-persistent \acr{edcs} and attention, 2) non-persistent \acr{edcs} and corticosteroids, and 3) corticosteroids and attention, using the parametric g-formula and average contrasts, in children of a large prospective birth cohort in Europe.