### 2023-11-16 output/paper/figures.qmd
Modified how paper is rendered, moving away from targets for that and simply calling quarto render on SI, tables, figures, and manuscript, and including sections of manuscript in manuscript.qmd. Still, cross ref does not work like this since they are rendered independently.

---
title: "Figures"

format:
  pdf:
    pdf-engine: lualatex
    number-sections: true
    toc: false
    include-in-header:
      text: |
        \usepackage{lineno}
        \usepackage{rotating}
        \usepackage{float}
        \usepackage{typearea}
    include-before-body:
      text: |
        \linenumbers

link-citations: true
linkcolor: blue

editor_options: 
  chunk_output_type: console

execute:
  echo: false
  warning: false
  error: false
  message: false
---

```{r}
here::i_am("output/paper/figures.qmd")
```

```{r vars}
source(here::here("code/dictionaries.R"))
source(here::here("code/data.R"))
source(here::here("code/proc_res.R"))

is_hpc <- FALSE
store <- ifelse(
  is_hpc, 
  paste0(
    "/PROJECTES/HELIX_OMICS/DATA_PREVIOUS_MIGRATION/",
    "lorenzoF/data/data_paper3/_targets/_targets_res"
  ), 
  "~/mounts/rstudioserver/data/data_paper3/_targets/_targets_res"
)

id_var <- "HelixID"
num_digits_est = 3
num_digits_sig = 2
params <- params(is_hpc = is_hpc)
threshold_within_lod <- params$steps$rq2$preproc_outcome$llodq$threshold_within
threshold_overall_lod <- params$steps$rq2$preproc_outcome$llodq$threshold_overall
threshold_within_nan <- params$steps$rq2$preproc_outcome$missings$threshold_within
threshold_overall_nan <- params$steps$rq2$preproc_outcome$missings$threshold_overall
knn_k <- params$steps$rq2$preproc_outcome$missings$k
```

- Figure titles and legends should be submitted as single paragraph.

# Main results

## Marginal comparisons

```{r}
targets::tar_load(
  dplyr::matches("marginal_comparisons_(\\d)$"),
  store = store
)
```

```{r marginal-1}
#| label: fig-marginal-1
#| fig-cap: Average comparisons
#| fig-width: 6
#| fig-height: 7

tidy_res_meffects(
  df = marginal_comparisons_1,
  sa_var = NULL,
  which_res = "comparisons",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$plot
```

```{r marginal-2}
#| label: fig-marginal-2
#| fig-cap: Average comparisons
#| fig-width: 9
#| fig-height: 10

tidy_res_meffects(
  df = marginal_comparisons_2,
  sa_var = NULL,
  which_res = "comparisons",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$plot
```

```{r marginal-3}
#| label: fig-marginal-3
#| fig-cap: Average comparisons
#| fig-width: 5
#| fig-height: 5

tidy_res_meffects(
  df = marginal_comparisons_3,
  sa_var = NULL,
  which_res = "comparisons",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$plot
```