### 2023-11-21 output/paper/sections_old/_introduction.qmd
Added presentation for group meeting regarding paper 3. Moved individual sections manuscript directly in manuscript. Added include of SI and tables and figures in manuscript for cross referencing.

The prevalence of several neurodevelopmental disorders has increased in the pediatric population [@GrandjeanLandrigan:2014], and multiple environmental pollutants play a role in the increased rates of these disorders [@RamirezGalvez-OntiverosGonzalez-Domenech:2022]. \acr{edcs}, chemicals that are capable of interfering with the endocrine system, have been shown to have a role in childhood neurodevelopment [@Braun:2017]. Although both pregnancy and early childhood are crucial stages of (neuro)development, most of the available literature is focused on the effects of prenatal exposure to \acr{edcs} on child neurodevelopment [@RamirezGalvez-OntiverosGonzalez-Domenech:2022]. Moreover, little is known about the biological mechanisms of action [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].

\acr{opps} might have a deleterious effect on neurodevelopment, although the few studies conducted on children and using a biological matrix for exposure assessment suffered from a series of limitations, including a small sample size and few \acr{opp} metabolites measured [@RamirezGalvez-OntiverosGonzalez-Domenech:2022]. Phthalates and their metabolites have been associated to several adverse effects on neurodevelopment [@RamirezGalvez-OntiverosGonzalez-Domenech:2022]. Similarly, these studies focused on few phthalates metabolites and were conducted in small populations. On the other hand, the effects of exposure to \acr{bpa} on cognitive functions are still unclear [@RamirezGalvez-OntiverosGonzalez-Domenech:2022].

Another major limitation of these epidemiological studies, especially when non-persistent chemicals are analyzed, is the reliance on single-spot urine samples for exposure assessment, which might not be representative of long-term exposures.

Something about corticosteroids...

In the present study, we estimated associations between 1) non-persistent \acr{edcs} and attention, 2) non-persistent \acr{edcs} and corticosteroids, and 3) corticosteroids and attention, using the parametric g-formula and average contrasts, in children of a large prospective birth cohort in Europe.