### 2023-11-21 output/presentations/cep_internal_231122/presentation.qmd

---
author: Lorenzo Fabbri
title: "Estimating associations between childhood exposure to non-persistent endocrine disruptors, corticosteroids, and neurodevelopment: A study based on the parametric g-formula"

format:
  revealjs:
    embed-resources: true
    slide-number: true
    width: 1600
    height: 900
    theme: simple
    center: true

editor_options: 
  chunk_output_type: console
execute:
  echo: false
  warning: false
  error: false
  message: false
---

## Recap

* We want to estimate the following *effects* (in children from HELIX):
  * Non-persistent EDCs and neurodevelopment (using attention as *proxy*).
  * Non-persistent EDCs and corticosteroids.
  * Corticosteroids and neurodevelopment.
* For attention we used the HIT-RT SE from the Attention Network Test.
* We did not consider individual metabolites but sums and ratios.
* We want to do so using the parametric g-formula and average contrasts.

## The pipeline

1. Estimation of balancing weights: `exposure ~ covariates`.
1. G-computation:
    1. Fit a **weighted outcome model**: `outcome ~ exposure + covariates`.
    1. Create two copies of the original dataset, with the exposure **shifted**.
    1. Predict the outcomes in the two **counterfactual** datasets.
    1. Estimate the **estimand** $\phi^\Delta = \mathbb{E} [Y^{d_1}] - \mathbb{E} [Y^{d_2}]$.

In this case:

* Cohort-specific shifting of the exposure. $d_1$ is a deterministic function setting the exposure to its 90th percentile, $d_2$ to its 10th percentile.

We also looked into **effect-modification** by sex (weights estimated separately, interaction in outcome model, estimation of marginal effects taking this into account).

## Difficulties {.smaller}

It looks simple, but lots of literature searches:

* DAGs for three different research questions.
* New data (corticosteroids), and still awaiting answers, and ratios difficult to interpret.
* Estimation of balancing weights for **ratios** is (apparently) still an open question.
* Usage of more **flexible modeling** approaches to estimate the ADRF (natural cubic splines).
* Careful modeling of **ratios**; it's not that simple:
    * `log (num / den) ~ exposure + log den + covariates`.
    * `outcome ~ log (num / den)`.
* Creatinine coming from different sources.
* Proper modeling of creatinine to take into account dilution effects.
* Careful consideration of positivity violations when shifting the exposures (cohort-specific and smaller percentiles).
* Computation of **robust** standard errors for inference using `cohort` as clusters.
* Careful usage of tools to improve **replicability** across the entire project (from data loading to paper writing).

## Results

See files...