### 2023-12-31 docs/bib_edc_cort_neuro.qmd
Finished doc with references.

---
title: "Some references on EDCs, glucocorticosteroids, and neurodevelopment"

format:
  pdf:
    pdf-engine: pdflatex
    number-sections: true
    number-depth: 2
    toc: true
    toc-depth: 2

bibliography: ../output/paper/bibliography.bib

link-citations: true
linkcolor: blue

execute:
  echo: false
  warning: false
  error: false
  message: false

editor_options:
  chunk_output_type: console
---

``` {r template}
#| echo: false
#| eval: false

## @

- Title: **.
- Exposures: .
- Medium exposure assessment: .
- Timing exposure assessment: .
- Outcomes: .
- Medium outcome assessment: .
- Timing outcome assessment: .
- Results:
    - .
- Sex-specific results: .
```

# EDCs and neurodevelopment

## @BouchardBellingerWright:2010

- Title: *Attention-Deficit/Hyperactivity Disorder and Urinary Metabolites of Organophosphate Pesticides*.
- Exposures: dialkylphosphate metabolites (biomarkers of organophosphate pesticide exposure).
- Medium exposure assessment: spot urine samples.
- Timing exposure assessment: ages 8-15.
- Outcomes: ADHD diagnostic status.
- Timing outcome assessment: ages 8-15.
- Results:
  - Exposure to OP pesticides might contribute to ADHD.
- Sex-specific results: NA.

## @CartierWarembourgLeManer-Idrissi:2016

- Title: *Organophosphate Insecticide Metabolites in Prenatal and Childhood Urine Samples and Intelligence Scores at 6 Years of Age: Results from the Mother–Child PELAGIE Cohort*.
- Exposures: dialkylphosphate metabolites (biomarkers of organophosphate pesticide exposure).
- Medium exposure assessment: spot urine samples.
- Timing exposure assessment: age 6.
- Outcomes: Wechsler Intelligence Scale for Children.
- Timing outcome assessment: age 6.
- Results:
  - No association between total DAP metabolites and WISC scores.
  - WISC working memory score inversely associated with levels of DE metabolites.
- Sex-specific results: NA.

## @FurlongHerringBuckley:2017

- Title: *Prenatal exposure to organophosphorus pesticides and childhood neurodevelopmental phenotypes*.
- Exposures: organophosphorus pesticides. Total diethylphosphate and total dimethylphosphate.
- Medium exposure assessment: spot urine samples.
- Timing exposure assessment: prenatal between 25 and 40 weeks of gestation.
- Outcomes: behavior, executive functioning, and IQ.
- Timing outcome assessment: ages 6–9.
- Results:
  - Dimethylphosphate metabolites negatively associated with Internalizing factor scores.
  - Dimethylphosphate metabolites positively associated with Executive Functioning factor scores.
  - Diethylphosphate metabolites negatively associated with Working Memory Index.
- Sex-specific results: no interactions by sex.

## @Gonzalez-AlzagaHernandezRodriguez-Barranco:2015

- Title: *Pre- and postnatal exposures to pesticides and neurodevelopmental effects in children living in agricultural communities from South-Eastern Spain*.
- Exposures: dialkylphosphate metabolites (biomarkers of organophosphate pesticide exposure).
- Medium exposure assessment: urine samples.
- Timing exposure assessment: ages 6-11.
- Outcomes: Wechsler Intelligence Scale for Children.
- Timing outcome assessment: ages 6-11.
- Results:
  - DAP levels inversely associated with performance on intelligence quotient and verbal comprehension domain.
- Sex-specific results: stronger associations in males.

## @HuangChenSu:2015

- Title: *Association of Exposure to Endocrine-Disrupting Chemicals During Adolescence With Attention-Deficit/Hyperactivity Disorder–Related Behaviors*.
- Exposures: seven metabolite of phthalate esters.
- Medium exposure assessment: urine samples.
- Timing exposure assessment: prenatal and at ages 2, 5, 8, and 11.
- Outcomes: Bayley and Wechsler tests for assessing neurocognitive functions and intelligence (IQ).
- Timing outcome assessment: ages 2, 5, 8, and 11.
- Results:
  - Children’s phthalate exposure (MEOHP and sum of DEHP metabolites) inversely associated with cognitive development.
- Sex-specific results: NA.

## @HuangTsaiChen:2017

- Title: *Intellectual evaluation of children exposed to phthalate-tainted products after the 2011 Taiwan phthalate episode*.
- Exposures: 5 phthalate metabolites.
- Medium exposure assessment: first-morning urine samples.
- Timing exposure assessment: ages 3-12.
- Outcomes: Wechsler tests for assessing the children's intelligence quotient.
- Timing outcome assessment: ages 3-12.
- Results:
  - MEOHP, MnBP, and MiBP inversely associated with verbal intelligence.
- Sex-specific results: NA.

## @KimHongShin:2017

- Title: *The effects of maternal and children phthalate exposure on the neurocognitive function of 6-year-old children*.
- Exposures: phthalate metabolites.
- Medium exposure assessment: urine samples.
- Timing exposure assessment: age 6
- Outcomes: IQ scores and continuous performance test variables.
- Timing outcome assessment: age 6.
- Results:
  - DEHP metabolites (including MEHHP and MEOHP) inversely associated with intelligence, attention, and response time variability.
- Sex-specific results: NA.

## @LiZhangKuang:2018

- Title: *Relationship between bisphenol A exposure and attention-deficit/ hyperactivity disorder: A case-control study for primary school children in Guangzhou, China*.
- Exposures: bisphenol-A.
- Medium exposure assessment: spot urine samples.
- Timing exposure assessment: ages 6-12.
- Outcomes: ADHD.
- Timing outcome assessment: ages 6-12.
- Results:
  - Positive association between BPA levels and odds of ADHD.
- Sex-specific results: stronger associations in males.

## @OhKimKannan:2023

- Title: *Early childhood exposure to environmental phenols and parabens, phthalates, organophosphate pesticides, and trace elements in association with attention de cit hyperactivity disorder (ADHD) symptoms in the CHARGE study*.
- Exposures: phenols/parabens, phthalates, OP pesticides, trace elements.
- Medium exposure assessment: spot urine samples.
- Timing exposure assessment: ages 2-5.
- Outcomes: ADHD symptoms.
- Timing outcome assessment: ages 2-5.
- Results:
    - Positive associations of MECPP and MEHHP with Hyperactivity subscale scores and hyperactivity/impulsivity subdomain scores.
    - Positive associations of MNBP with hyperactivity/impulsivity subdomain scores.
    - Inverse associations of TCS with Hyperactivity subscale scores.
    - Inverse associations of TCS with inattention subdomain scores.
    - Positive associations of mixture of phthalates with scores of the ABC Hyperactivity subscale and hyperactivity/impulsivity subdomain.
      - Drivers by subdomain, hyperactivity/impulsivity subdomain: MHPP, MEHHP, and MBZP.
    - Positive associations of mixture of phenols/parabens with inattention subdomain score.
      - Drivers by subdomain, inattention subdomain: MEPB, PRPB, and OHMEP.
- Sex-specific results: association of mixture of phthalates with hyperactivity-related subscale and subdomain significant in males only.

## @Rodriguez-CarrilloMustielesPerez-Lobato:2019

- Title: *Bisphenol A and cognitive function in school-age boys: Is BPA predominantly related to behavior?*.
- Exposures: bisphenol-A.
- Medium exposure assessment: spot urine samples.
- Timing exposure assessment: ages 9-11.
- Outcomes: comprehensive neuropsychological test battery.
- Timing outcome assessment: ages 9-11.
- Results:
  - No consistent association between BPA levels and cognitive abilities.
  - Possible inverse association between BPA levels and working memory.
- Sex-specific results: assessment only in males.

## @ShoaffCoullWeuve:2020

- Title: *Association of Exposure to Endocrine-Disrupting Chemicals During Adolescence With Attention-Deficit/Hyperactivity Disorder–Related Behaviors*.
- Exposures: phthalates, parabens, phenols, and triclocarban.
- Medium exposure assessment: urine samples.
- Timing exposure assessment: age 15.3
- Outcomes: Conners Attention Deficit Scale and the Behavior Assessment System for Children.
- Timing outcome assessment: age 15.3
- Results:
  - Exposure to antiandrogenic phthalates increases the risk of ADHD-related behavior problems.
- Sex-specific results: stronger associations in males.

## @TewarAuingerBraun:2016

- Title: *Association of Bisphenol A exposure and Attention-Deficit/Hyperactivity Disorder in a national sample of U.S. children*.
- Exposures: bisphenol-A.
- Medium exposure assessment: spot urine samples.
- Timing exposure assessment: ages 8-15.
- Outcomes: presence of ADHD in the past year.
- Timing outcome assessment: ages 8-15.
- Results:
  - Positive association between BPA levels and ADHD.
- Sex-specific results: stronger associations in males.

## @VilmandBeckBilenberg:2023

- Title: *Prenatal and current phthalate exposure and cognitive development in 7-year-old children from the Odense child cohort*.
- Exposures: phthalates metabolites.
- Medium exposure assessment: spot urine samples.
- Timing exposure assessment: 7 years.
- Outcomes: IQ (subtests from Wechsler Intelligence Scale for Children).
- Timing outcome assessment: 7 years.
- Results:
    - Inverse associations of MCPP, sum of DnHxPm, sum of DiDPm, and sum of DiNPm with IQ points.
- Sex-specific results: no evidence of effect modification by sex.

## @YuDuChiou:2016

- Title: *Increased risk of attention-deficit/hyperactivity disorder associated with exposure to organophosphate pesticide in Taiwanese children*.
- Exposures: dialkylphosphate metabolites (biomarkers of organophosphate pesticide exposure).
- Medium exposure assessment: urine samples.
- Timing exposure assessment: ages 4-15.
- Outcomes: ADHD.
- Timing outcome assessment: ages 4-15.
- Results:
  - Dose–response relationship between urinary concentrations of dimethylphosphate and ADHD.
- Sex-specific results: NA.

---

# EDCs and glucocorticosteroids

## @KimLeeMoon:2018

- Title: *Association of phthalate exposures with urinary free cortisol and 8-hydroxy-2'-deoxyguanosine in early childhood*.
- Exposures: four phthalate metabolites.
- Medium exposure assessment: spot urine samples.
- Timing exposure assessment: prenatal and at ages 0, 3, 9, 12, 15 months.
- Outcomes: free cortisol.
- Medium outcome assessment: spot urine samples.
- Timing outcome assessment: prenatal and at ages 0, 3, 9, 12, 15 months.
- Results:
    - Positive associations between MEHHP, MEOHP, MiBP, and MnBP, with free cortisol.
- Sex-specific results: no evidence of modification by sex.

## @SearsLiuLanphear:2023

- Title: *Evaluating mixtures of urinary phthalate metabolites and serum per-/polyfluoroalkyl substances in relation to adolescent hair cortisol: The HOME Study*.
- Exposures: phthalate metabolites.
- Medium exposure assessment: spot urine samples.
- Timing exposure assessment: ages 1-5 and 8.
- Outcomes: cortisol.
- Medium outcome assessment: hair.
- Timing outcome assessment: age 12.
- Results:
    - Positive association between mixture of phthalates metabolites measured in childhood and hair cortisol measured in adolescence. Driven by MEP, MiBP, MBzP.
- Sex-specific results: no evidence of modification by sex.

## @SunLiJin:2018

- Title: *Associations between repeated measures of maternal urinary phthalate metabolites during pregnancy and cord blood glucocorticoids*.
- Exposures: phthalate metabolites.
- Medium exposure assessment: spot urine samples.
- Timing exposure assessment: 14, 24, and 36 weeks of gestation.
- Outcomes: cortisol and cortisone.
- Medium outcome assessment: cord blood.
- Timing outcome assessment: 14, 24, and 36 weeks of gestation.
- Results:
    - Positive association of MEHP 3rd trimester with cortisol.
    - Inverse association of MECPP and MEOHP 3rd trimester with cortisone.
    - Positive association of MBzP 1st trimester with cortisol/cortisone ration.
- Sex-specific results: evidence of modification by sex, but not significant.
    - Females: positive associations with glucocorticoids (MEHP and MBzP 1st trimester, and DEHP 3rd trimester).
    - Males: inverse associations with glucocorticoids (MECPP, MEHHP, and MEOHP 3rd trimester).

---

# Glucocorticosteroids and neurodevelopment

## @LupienMcEwenGunnar:2009

- Title: *Effects of stress throughout the lifespan on the brain, behaviour and cognition*.

The HPA axis, which can be activated by stress, is responsible for the production of glucocorticoids. The brain, and its proper functioning, is a potential target, due to the presence of receptors for these hormones.
Glucocorticoids are necessary for brain maturation, although their under- or over-production might interfere with its normal development and ultimately lead to long-term impaired functioning.

---

# References {.unnumbered}

::: {#refs}
:::