### 2024-02-06 output/paper/authors.qmd
Modified almost all text based on MV. Added authors and roles.

---
title: "Childhood exposure to non-persistent endocrine disruptors, glucocorticosteroids, and neurodevelopment: A study based on the parametric g-formula"

metadata-files:
  - ../../docs/authors.yml

format:
  html:
    embed-resources: true

link-citations: true
linkcolor: blue
---