### 2024-02-12 output/paper/authors.qmd
Switching to new syntax acr. Adding comments from MV v1.

---
title: "Childhood exposure to non-persistent endocrine disruptors, glucocorticosteroids, and neurodevelopment: A study based on the parametric g-formula"

metadata-files:
  - ../../docs/authors.yml

format:
  html:
    embed-resources: true
    template-partials:
      - title-metadata.html

link-citations: true
linkcolor: blue
---