### 2024-02-13 output/paper/authors.qmd
Finishing MV v1. Working on tables authors.

---
title: "Childhood exposure to non-persistent endocrine disruptors, glucocorticosteroids, and neurodevelopment: A study based on the parametric g-formula"

metadata-files:
  - ../../docs/authors.yml

format:
  html:
    embed-resources: true

link-citations: true
linkcolor: blue

editor_options:
  chunk_output_type: console
---

```{r}
authors <- yaml::read_yaml("../../docs/authors.yml")

proc_one <- function(df, idx) {
  tmp <- df$authors[[idx]]

  affiliations <- unname(unlist(tmp$affiliations))
  full_names <- tibble::tibble(unlist(df$affiliations))
  stringr::str_replace_all()

  ret <- tibble::tribble(
    ~name, ~affiliations, ~role,
    tmp$name, paste0(unname(unlist(tmp$affiliations)), collapse = "; "), paste0(tmp$role, collapse = "; ")
  )
}
```