### 2024-03-08 output/paper/_tables.qmd

```{r}
here::i_am("output/paper/_tables.qmd")
```

# Tables for descriptive data

## Study populations

::: {#tbl-pop-desc}
```{r pop-desc}
targets::tar_load("table1", store = store)
table1
```
**Participant characteristics (HELIX subcohort; 2013-2016).**
:::

## Endocrine disruptors

::: {#tbl-edc-desc}
```{r edc-desc}
targets::tar_load("desc_chems_mets", store = store)
desc_chems_mets$desc_chems
```
**Participants endocrine disruptors concentrations expressed in $\mu$grams/L (HELIX subcohort; 2013-2016).**
:::

## Glucocorticosteroids

::: {#tbl-met-new-desc}
```{r met-new-desc}
metabolites <- load_steroids()
scores <- create_steroid_scores(dat = metabolites$metabolome) |>
  tidylog::select(
    HelixID,
    cortisol_production, cortisone_production,
    corticosterone_production
  )
scores <- myphd::extract_cohort(
  dat = scores,
  id_var = "HelixID",
  st = 1, en = 3
) |>
  tidylog::select(-HelixID) |>
  tidylog::mutate(
    cohort = dplyr::case_when(
      cohort == "EDE" ~ "EDEN",
      cohort == "KAN" ~ "KANC",
      cohort == "MOB" ~ "MOBA",
      cohort == "RHE" ~ "RHEA",
      cohort == "SAB" ~ "INMA",
      .default = cohort
    )
  ) |>
  tidylog::rename(
    "cortisol production" = "cortisol_production",
    "cortisone production" = "cortisone_production",
    "corticosterone production" = "corticosterone_production"
  )

# Entire population (1004)
desc_sums_all <- scores |>
  dplyr::select(-cohort) |>
  gtsummary::tbl_summary(
    statistic = list(
      gtsummary::all_continuous() ~ c(
        "{median} ({p25}, {p75}); {N_miss} ({p_miss})"
      )
    ),
    missing = "no",
    digits = dplyr::everything() ~ 1
  )

# Intersection with HELIX subcohort (976)
targets::tar_load(
  dplyr::contains("load"),
  store = "~/mounts/rstudioserver/data/data_paper3/_targets/_targetsRQ2/"
)
desc_sums_intersection <- create_steroid_scores(rq2_load_dat$outcome) |>
  dplyr::select(
    cortisol_production,
    cortisone_production,
    corticosterone_production
  ) |>
  tidylog::rename(
    "cortisol production" = "cortisol_production",
    "cortisone production" = "cortisone_production",
    "corticosterone production" = "corticosterone_production"
  ) |>
  gtsummary::tbl_summary(
    statistic = list(
      gtsummary::all_continuous() ~ c(
        "{median} ({p25}, {p75}); {N_miss} ({p_miss})"
      )
    ),
    missing = "no",
    digits = dplyr::everything() ~ 1
  )

gtsummary::tbl_merge(
  tbls = list(desc_sums_all, desc_sums_intersection)
) |>
  gtsummary::modify_spanning_header(
    dplyr::everything() ~ NA_character_
  ) |>
  gtsummary::as_gt() |>
  gt::tab_footnote(
    footnote = "Measurements available for the HELIX subcohort.",
    locations = gt::cells_column_labels(8)
  ) |>
  gt::opt_footnote_marks(
    marks = "letters"
  )
```
**Participants derived glucocorticosteroids concentrations expressed in ng/ml (HELIX subcohort; 2013-2016).**
:::

# Tables for other analyses

## Marginal hypotheses for effect modification

```{r}
targets::tar_load(
  dplyr::matches("marginal_hypothesis_(\\d)SA$"),
  store = store
)
edcs <- myphd::edcs_information() |>
  dplyr::filter(!chem_id %in% c("dedtp", "dmdtp")) |>
  dplyr::select(chem_id, short_name)
```

::: {#tbl-hypothesis-1and2}
```{r hypothesis-1and2}
t1 <- tidy_res_meffects(
  df = marginal_hypothesis_1SA,
  sa_var = "e3sex",
  outcome = NULL,
  which_res = "hypothesis",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$table
t2 <- tidy_res_meffects(
  df = marginal_hypothesis_2SA,
  sa_var = "e3sex",
  outcome = NULL,
  which_res = "hypothesis",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$table

dplyr::bind_cols(t1$`_data`, t2$`_data`) |>
  dplyr::select(-`class...4`, -`variable...5`) |>
  dplyr::rename(
    class = "class...1",
    variable = "variable...2"
  ) |>
  dplyr::left_join(
    edcs,
    by = c("variable" = "chem_id")
  ) |>
  dplyr::select(-variable) |>
  dplyr::rename(variable = "short_name") |>
  gt::gt(
    rowname_col = "variable",
    groupname_col = "class"
  ) |>
  gt::tab_style(
    style = gt::cell_text(
      weight = "bold"
    ),
    locations = gt::cells_row_groups()
  ) |>
  gt::tab_footnote(
    footnote = "Estimate and 95% CI.",
    locations = gt::cells_column_labels()
  ) |>
  gt::opt_footnote_marks(
    marks = "letters"
  )
```
**Pairwise differences between marginal contrasts of males and females, for the effect of a decrease from the 90th to the 10th percentile of {{< acr edcs first_use=true >}} on {{< acr hitrtse first_use=true >}}, expressed in {{< acr mss >}}, and on the glucocorticosteroids, expressed in ng/ml (HELIX subcohort; 2013-2016).**
:::

::: {#tbl-hypothesis-3}
```{r hypothesis-3}
tidy_res_meffects(
  df = marginal_hypothesis_3SA,
  sa_var = "e3sex",
  outcome = NULL,
  which_res = "hypothesis",
  num_digits_est = num_digits_est,
  num_digits_sig = num_digits_sig
)$table
```
**Pairwise differences between marginal contrasts of males and females, for the effect of a decrease from the 90th to the 10th percentile of the glucocorticosteroids on {{< acr hitrtse first_use=true >}} expressed in {{< acr mss >}} (HELIX subcohort; 2013-2016).**
:::